\documentclass[twocolumn,
  showpacs,showkeys,prd,superscriptaddress]{revtex4-1}

\usepackage{amsmath,amsthm,latexsym,amssymb,amsfonts,epsfig}
\usepackage{xcolor}
\usepackage{ulem}
%\usepackage{authblk}
\usepackage[%
  colorlinks=true,
  urlcolor=blue,
  linkcolor=blue,
  citecolor=blue
]{hyperref}
\usepackage{etoolbox}
\usepackage{breqn}
\makeatletter
\let\cat@comma@active\@empty
\makeatother

%------------------
%--------- Definitions
%------------------
%% \def\be{\begin{equation}}
%% \def\ee{\end{equation}}
%% \def\ba{\begin{eqnarray}}
%% \def\ea{\end{eqnarray}}
\def\md{{\mathrm{d}}}
%\def\mx{{\mathrm{x}}}
\def\my{{\mathrm{y}}}
\def\mz{{\mathrm{z}}}
\def\l{\left}
\def\r{\right}
\def\ll{\left\{}
\def\rl{\right\}}
\def\lp{\left(}
\def\rp{\right)}
\def\lc{\left[}
\def\rc{\right]}
%% \renewcommand{\{}{\left\{}
%% \renewcommand{\}}{\right\}}
%% \renewcommand{\[}{\left[}
%% \renewcommand{\]}{\right]}
%% \renewcommand{\(}{\left(}
%% \renewcommand{\)}{\right)}
\def\la{\left\langle}
\def\ra{\right\rangle} 
%% \DeclareMathOperator{\det}{det}
%% \newcommand\tor[3]{\ensuremath{T_{#1}{}^{#2}{}_{#3}}}
%% %\renewcommand\Authands{ and }
%% \newcommand{\out}[1]{{\color{red} \sout{#1}}}
\newcommand{\hl}[1]{{\color{red} \textbf{#1}}}
%% \newcommand{\pro}[1]{{\color{blue} #1}}

\input{Def-article.tex}


\hypersetup{%
  pdftitle={Purely Affine Gravity},
  pdfauthor={Aureliano Skirzewski,}{Oscar Castillo-Felisola},
  pdfkeywords={Affine Gravity,} {Torsion,} {Generalised Gravity.},
  pdflang={English}
}



%------------------
%--------- Document
%------------------
\begin{document}

\title{Purely affine gravity}


\author{Aureliano \surname{Skirzewski}}
\email[Corresponding Author: ]{skirz@ula.ve}
\affiliation{Centro de F\'\i sica Fundamental,  Universidad de los Andes, 5101 M\'erida, Venezuela.}

\author{Oscar \surname{Castillo-Felisola}}
\email{o.castillo.felisola@gmail.com}
\affiliation{Centro Cient\'\i fico Tecnol\'ogico de Valpara\'\i so, Chile.}
\affiliation{Departamento de F\'\i sica, Universidad T\'{e}cnica Federico Santa Mar\'\i a, Casilla 110-V, Valpara\'\i so, Chile.}

%--------- Abstract
\begin{abstract}
  We develop a topological theory of gravity with torsion where metric has a dynamical rather than a kinematical origin. This approach towards gravity resembles pre-geometrical approaches in which a fundamental metric is not assumed even though the  affine connection gives place to a local inertial structure, which reminds us of Mach's principle that assumes the inertial forces, subject to locally Minkowskian metric structures, should have dynamical origin. Additionally a Newtonian like gravitational force is obtained in the perturbative limit of the theory.
\end{abstract}

\pacs{04.50.-h,04.62.+v,11.25.Mj,11.25.-w}
\keywords{Affine Gravity, Torsion, Generalised Gravity.}

\maketitle

\section{Introduction}

General Relativity (GR) has proven to be the most successful theory of gravity.  However,  it is not as successful as we may wish. Part of the problem with GR is that the standard quantization procedure cannot be applied  properly on it. Moreover, not only it is not renormalizable, but there are  problems with the choice of variables to be quantized and the choice of the Hilbert space to be used. 

Although there is nothing wrong about metric spacetimes,  certainly it is an issue to sum over all possible field configurations of the metric, as this would imply summing Euclidean and Minkowski like contributions to the transition amplitudes on equal terms. Additionally, we might also consider the difficulties of  quantizing  non polynomial functions of the three-dimensional metric that appears in the Hamiltonian in an ADM formulation of GR, such as the square root of the metric's determinant or the Ricci scalar.

In order to address some of these issues, several approaches have been designed that use the connection as a fundamental field. For instance, in the context of Cartan formulations of gravity, letting the metric background become flat we can rewrite Einstein-Hilbert's Lagrangian as a function of the torsion field. This approach is known as Teleparallel Gravity and is equivalent to GR in spite of taking torsion as the fundamental gravitational field.

Furthermore, another alternative description of GR developed initially by Ashtekar uses the spin connection as the fundamental field and the frame field turns out to be its canonically conjugated momentum. In the context of Loop Quantum Gravity (LQG), using Ashtekar variables, a successful quantization program has been achieved. Some  strengths of this quantization program lies within a theorem by Hanno Sahlman that states the only diffeomorphisms invariant Hilbert space that supports the Heisenberg algebra, for the connection and its associated momentum, is the one of Loop Quantum Gravity. In spite of its success, LQG has not advanced enough to conclude that its low energy effective description is GR. Currently, there is no clue about the LQG effective description at other scales,  nor its continuum spacetime limit either. Therefore, we cannot conclude that the search for a fundamental theory of gravitational interactions has ended. On the contrary, there are many alternatives to the usual metric description of Gravity and they all must be tested against experiments and observations.

In this article we study a diffeomorphism invariant ``toy'' model  consisting  solely of an affine connection  (with torsion), and the strict commitment to formulate a power counting renormalizable action.  We find the condition on the background metric to be exactly zero to be appealing because the use of a background metric such as in teleparallel theories of gravity would in principle break background independence. We expect  this strategy may serve to overcome  the uniqueness theorem about diffeomorphism invariant theories of connections, since we  have no fundamental metric field to quantize. The earliest model (that we know of) that argues  a description of gravitational interaction in terms of connections as fundamental fields  was presented by Eddington~\cite{Eddington1923math}, for an spacetime with positive cosmological constant. He proposed the square root of the determinant of the Ricci tensor as the gravitational Lagrangian. His aim was not to solve the problems of quantum gravity,  others have emphasized the character of GR as a gauge theory in order to address the issues of quantization and regularization, as in LQG. Among other, the work by  Krasnov studies a BF model that is conceived to {\bfseries \color{blue} TO BE CONTINUED...}

We aim to describe certain aspects of the gravitational interaction in four dimensions. To this end we have studied the non-relativistic limit of the parallel transport equations of motion of a particle moving on a flat, static, homogeneous and isotropic background.

%% \vspace*{5mm}

%% \hrule

%% \vspace*{5mm}

%% General Relativity (GR) is nowadays the accepted theory of gravitation. The gravitational effects are interpreted as a geometrical influence due to the curvature of the spacetime. In principle, the spacetime is a (pseudo-) Riemannian manifold whose metric tensor is the fundamental field in the description of the gravity.

%% More generally, the spacetime might be associated with a topological space, $\mathcal{M}$, with less restrictive structure than a (pseudo-) Riemannian geometry  by choosing a connection non-compatible with the metric, in the sense that $\nabla g\neq 0$. It is  still possible to define a connection associated with the parallel transport of vectors along curves on the spacetime manifold. Interestingly, this general connection cannot be written in terms of the metric, but decomposes into the usual Levi-Civita connection plus a contribution from the torsion tensor,
%% \begin{equation}
%%   \Gamma^{\mu}{}_{\rho\sigma} = \hat{\Gamma}^{\mu}{}_{\rho\sigma} + T^{\mu}{}_{\rho\sigma},
%% \end{equation}
%% where the hatted objects represent the the torsion-free part (in the case of the connection it is just the Levi-Civita connection).

%% Therefore, another possible formulation of a gravitational theory within the context of a general connection would have an action like,
%% \begin{equation}
%%   S[g,\Ga]=\int\dn{n}{x}\sqrt{g}\; g^{\mu\nu}\Ri_{\mu\nu}\(\Ga,\partial\Ga\).
%% \end{equation}
%% This theory is known as the metric-affine gravity. It should be noticed that the above generalization of gravity is ``minimal'' in the sense that the new tensor (the torsion) enters only through the minimal coupling, while more general actions admit quadratic terms in the torsion.

%% \vspace*{5mm}

%% \hrule

%% \vspace*{5mm}

The aim of this work is to present a purely affine theory of gravity, where the symmetric tensor which acts like ``metric'' is a composite field defined in terms of the connection, \hl{Is it still TRUE?}
\begin{equation}
  g_{\mu\nu} = T^\sigma{}_{\rho\mu} T^\rho{}_{\sigma\nu}.
\end{equation}

The structure of the paper ...

%A gravitational theory can still be formulated in this non-metric spacetimes, and are known as \textit{affine gravity}.



%--------- Scattered Ideas
\section{Warming up: The three-dimensional case}

Formally, the curvature of a manifold is defined through the commutator of covariant derivatives under diffeomorphims, $\nabla_\mu$, but for general choice of the connection, $\Ga^\mu{}_{\nu\lambda}$, there is an extra contribution given by its antisymmetric part in the lower indices, $T^\mu{}_{\nu\lambda}=2\Ga^\mu{}_{[\nu\lambda]}$. Therefore, the commutator of the covariant derivatives yields,
\begin{equation}
  \comm{\nab{\mu}}{\nab{\nu}}\omega_\lambda = R_{\mu\nu}{}^\rho{}_\lambda \omega _\rho - T^\rho{}_{\mu\nu}\nab{\rho}\omega_\lambda.\label{curvdef}
\end{equation}
Note that $T^\rho{}_{\mu\nu}$ is a  tensor under diffeomorphisms.

In order to build topological invariants of density one, one can use the skew-symmetric Levi-\v{C}ivita tensor $\epsilon^{\mu_1\mu_2\dots\mu_n}$ in $n$-dimensional space(-time).

\begin{widetext}
  With these ingredients, in a three-dimensional space one can write an action 
  \begin{dmath}[compact, spread=2pt]
    \label{accion3d}
       S[\Gamma] =
      \int \dn{3}{x}  \Bigg\{R_{\mu_1\mu_2}{}^\rho{}_{\mu_3} T^\sigma{}_{\mu_4\mu_5}\sum_{\pi \in  \mathrm{Z}_5}C_\pi\delta_\rho^{\mu_{\pi(1)}} \delta_\sigma^{\mu_{\pi(2)}}\epsilon^{\mu_{\pi(3)}\mu_{\pi(4)}\mu_{\pi(5)}} 
       + T^\rho{}_{\mu_1\mu_2} T^\sigma{}_{\mu_3\mu_4}T^\tau{}_{\mu_5\mu_6} \sum_{\pi \in \mathrm{Z}_6}D_\pi      \delta_\rho^{\mu_{\pi(1)}} \delta_\sigma^{\mu_{\pi(2)}}\delta_\tau^{\mu_{\pi(3)}}\epsilon^{\mu_{\pi(4)}\mu_{\pi(5)}\mu_{\pi(6)}} 
       +T^\rho{}_{\mu_1\mu_2}\nabla_{\mu_3} 
       T^\sigma{}_{\mu_4\mu_5}\sum_{\pi \in \mathrm Z_5}E_\pi\delta_\rho^{\mu_{\pi(1)}} \delta_\sigma^{\mu_{\pi(2)}}\epsilon^{\mu_{\pi(3)}\mu_{\pi(4)}\mu_{\pi(5)}} \Bigg\}, 
  \end{dmath}
  where all possible permutations of $n$ elements $\pi\in\mathrm{Z}_n$ have been included in the sums with  different constants $C_\pi$, $D_\pi$ and $E_\pi$ for  permutation. 
\end{widetext}

The torsion field can be decomposed into invariant tensors respecting the symmetry,
\begin{equation}
  T^\sigma{}_{\mu\nu}=\epsilon_{\mu\nu\rho} T^{\sigma\rho}+A_{[\mu}\delta^\sigma{}_{\nu]},
\end{equation}
with a symmetric $T^{\sigma\rho}$ of density weight  $w =1$, and \mbox{$A_\mu=T^\nu{}_{\mu\nu}$} is the trace part of the more arbitrary $T^\sigma{}_{\mu\nu}$.

The action in Eq.~\eqref{accion3d} can be rewritten as
\begin{dmath}[compact, spread=2pt]
  \label{accion3dnew}
  S[\Gamma]=\int\dn{3}{x} \Bigg( 
  B_1R_{\mu\nu}{}^\mu{}_{\rho} T^{\nu\rho} 
  +B_2\epsilon^{\mu\nu\rho}R_{\mu\nu}{}^{\sigma}{}_\sigma A_\rho 
  +B_3 \epsilon^{\mu\nu\rho}R_{\mu\nu}{}^\lambda{}_\rho A_\lambda 
  +B_4 \det(T^{\mu\nu}) 
  +B_5 T^{\mu\nu}A_\mu A_\nu 
  +B_6 T^{\mu\nu}\nabla_\mu A_\nu
  +B_7\epsilon^{\mu\nu\rho}A_\mu\partial_\nu A_\rho
  + B_8\epsilon^{\mu\nu\rho}\Gamma^{\sigma}{}_{\mu\sigma}\partial_\nu\Gamma^{\tau}{}_{\rho\tau}
  +B_9\epsilon^{\mu\nu\lambda}\Big(\Gamma^{\sigma}{}_{\mu\rho}\partial_\nu\Gamma^{\rho}{}_{\lambda\sigma}
  +\frac{2}{3}\Gamma^{\tau}{}_{\mu\rho}\Gamma^{\rho}{}_{\nu\sigma}\Gamma^{\sigma}{}_{\lambda\tau}{}\Big)
   \Bigg).
\end{dmath}
where the coefficients $B_i$ are related with the original coefficients $C_i$, $D_i$ and $E_i$, and the additional $B_9$ term can be added in three dimensions leaving the action invariant under diffeomorphisms. The affine connection can be decomposed into its symmetric and antisymmetric parts, 
\begin{equation}
  \Gamma^\lambda{}_{\mu\rho}=\hat{\Gamma}^\lambda{}_{(\mu\rho)} + \epsilon_{\mu\rho\sigma}T^{\lambda\sigma} + A_{[\mu}\delta^\lambda{}_{\rho]},
\end{equation}
where  $\epsilon_{\mu\rho\sigma}$ has been introduced, and it is related to the skew symmetric $\epsilon^{\mu\rho\sigma}$ through the identity $\epsilon^{\lambda\mu\nu}\epsilon_{\rho\sigma\tau}=3!\delta^{\lambda}{}_{[\rho}\delta^\mu{}_{\sigma}\delta^{\nu}{}_{\tau]}$. Therefore, the curvature tensor can be expressed as 
\begin{dmath}[compact, spread=2pt]
  \label{RiemmanDecomposition}
  R_{\mu\nu}{}^\sigma{}_\rho=
  \hat{R}_{\mu\nu}{}^\sigma{}_\rho
  -2\epsilon_{\rho\alpha[\mu}\hat\nabla_{\nu]}T^{\sigma\alpha}
  +\partial_{[\mu}A_{\nu]}\delta^\sigma_\rho
  +\delta^\sigma_{[\mu}\hat\nabla_{\nu]}A_\rho
  +\epsilon_{\mu\nu\kappa}T^{\kappa\sigma}A_\rho
  -\delta^\sigma_{[\mu}\epsilon_{\nu]\rho\alpha}T^{\alpha\beta}A_\beta 
  +\frac{1}{2}\delta^\sigma_{[\mu}A_{\nu]}A_\rho
  -2\epsilon_{\alpha\beta[\mu}\epsilon_{\nu]\rho\delta}T^{\sigma\alpha}T^{\beta\delta},
\end{dmath}
where $ \hat\nabla_\rho$ and $\hat{R}_{\mu\nu}{}^\lambda{}_\rho$ are the covariant derivative and  curvature associated to the symmetric part of the connection. Notice that Bianchi identity, obtained as $\epsilon^{\mu\nu\lambda}\hat R_{\mu\nu}{}^\rho{}_\lambda=0$, leads us to the following
\begin{equation}
 \label{bianchi}
 \epsilon^{\mu\nu\rho} R_{\mu\nu}{}^\lambda{}_\rho = 4\hat\nabla_\rho T^{\rho\lambda}
+2\epsilon^{\mu\nu\lambda}\partial_\mu A_\nu-4T^{\lambda\rho}A_\rho. 
\end{equation}
Using the  Eqs.~\eqref{RiemmanDecomposition} and \eqref{bianchi} one can rewrite the action in Eq.~\eqref{accion3dnew} as
\begin{dmath}[compact, spread=2pt]
  \label{accion3dfinal}
  S[\Gamma]=\int\dn{3}{x} \Bigg( 
  (B_1+2B_9) \hat R_{\mu\nu}{}^\mu{}_{\rho} T^{\nu\rho} 
  +(B_2+B_9+B_9) \epsilon^{\mu\nu\rho}\hat R_{\mu\nu}{}^{\sigma}{}_\sigma A_\rho
  +(-6B_1+B_4-4B_9) \det(T^{\mu\nu}) 
  +(\frac{1}{2}B_1-4B_3+B_5+B_9 ) T^{\mu\nu}A_\mu A_\nu 
  +(B_1-4B_3+B_6+2B_9) T^{\mu\nu}\hat\nabla_\mu A_\nu
  +(2B_2+2B_3+B_7+B_9)\epsilon^{\mu\nu\rho}A_\mu\partial_\nu A_\rho
  +B_9\epsilon^{\mu\nu\lambda}\Big(\hat\Gamma^{\sigma}{}_{\mu\rho}\partial_\nu\hat\Gamma^{\rho}{}_{\lambda\sigma}
  +\frac{2}{3}\hat\Gamma^{\tau}{}_{\mu\rho}\hat\Gamma^{\rho}{}_{\nu\sigma}{}\hat\Gamma^{\sigma}{}_{\lambda\tau}{}\Big)
  + B_8\epsilon^{\mu\nu\rho}\hat\Gamma^{\sigma}{}_{\mu\sigma}\partial_\nu\hat\Gamma^{\tau}{}_{\rho\tau}
  +\partial_\alpha\Big( 4B_3T^{\alpha\mu} A_\mu 
  +B_9(2\Gamma^{[\delta}{}_{\delta\beta}T^{\alpha]\beta}-T^{\alpha\beta}A_\beta+\frac{1}{2}\epsilon^{\beta\alpha\eta}A_\eta\Gamma^{\delta}{}_{\delta\beta})
  +B_8\epsilon^{\beta\alpha\eta}A_\eta\hat\Gamma^{\delta}{}_{\delta\beta}\Big)
  \Bigg),
\end{dmath}
or after dropping the boundary terms and rename the coefficients,
\begin{dmath}[compact, spread=2pt]
  \label{3Dfull}
  S[\hat\Gamma,T,A] =
  \int \dn{3}{x} \bigg( 
  A_1\hat R_{\mu\nu}{}^{\mu}{}_\rho T^{\nu\rho} 
  +A_2\epsilon^{\mu\nu\rho}\hat R_{\mu\nu}{}^{\sigma}{}_\sigma A_\rho
  +A_3\epsilon^{\mu\nu\rho}A_\mu\partial_\nu A_\rho
  +A_4T^{\mu\nu}\hat\nabla_\mu A_\nu
  +A_5T^{\mu\nu}A_\mu A_\nu
  +A_6\det(T^{\mu\nu}) 
  +A_7\epsilon^{\mu\nu\lambda}\Big(\hat\Gamma^{\sigma}{}_{\mu\rho}\partial_\nu\hat\Gamma^{\rho}{}_{\lambda\sigma}
  +\frac{2}{3}\hat\Gamma^{\tau}{}_{\mu\rho}\hat\Gamma^{\rho}{}_{\nu\sigma}{}\hat\Gamma^{\sigma}{}_{\lambda\tau}{}\Big)
  + A_8\epsilon^{\mu\nu\rho}\hat\Gamma^{\sigma}{}_{\mu\sigma}\partial_\nu\hat\Gamma^{\tau}{}_{\rho\tau}
  \bigg).
\end{dmath}



%Poplawski...
Noticing that in  the first term, the variation respect to the Ricci tensor yields to $T^{\mu\nu}$, it can be argued that in a standard theory of gravity this tensor density corresponds to $\sqrt{g}g^{\mu\nu}$.
Therefore, Eq.~\eqref{accion3dfinal} reveals a one to one correspondence with General Relativity non minimally coupled to the $A_\mu$ field, \hl{I'd say that this result should be presented in terms of the action with coefficients C's... This result should be review!!!}
\begin{dmath}
  \label{accion3dGR}
  %\begin{split}
    S[g,\hat\Gamma,A] =
    \int\dn{3}{x} \Bigg\{\sqrt{g}\Big(B_1\hat R +\frac{1}{4}(4B_4-3B_1) -\frac{1}{4}\l( 2B_2-B_3\r) A_\mu A^\mu\Big) + \frac{1}{2}(2B_2-B_3)\epsilon^{\mu\nu\rho}A_\mu\hat R_{\nu\rho}{}^{\sigma}{}_\sigma+\frac{1}{4}(B_1+4B_2-2B_3)\epsilon^{\mu\nu\rho}A_\mu\partial_\nu A_\rho\Bigg\}
  %\end{split}
\end{dmath}
Thus, an interesting sector of the theory corresponds to the space of non-degenerated $T^{\mu\nu}$. 

As a matter of fact, the vector field and the ``gravitational" sector of the torsion disengage when $B_3=2B_2$

%% After these considerations we can conclude that the final expression for the action functional is then supposed to be expressed as a linear combination of all the invariants we can construct from $\hat{\Gamma}^\rho{}_{\mu\rho}$, $T^{\mu\nu}$ and $A_\mu$. Thus, the most general action should be 
%% \begin{equation}
%%   \begin{split}
%%     S[\hat{\Gamma}^\nu{}_{\mu\rho},T^{\mu\nu},A_\mu] = \int \dn{3}{x} & \bigg(
%%     C_1\hat{R}_{\mu\nu}{}^{\mu}{}_\rho T^{\nu\rho}\\
%%     & + \epsilon^{\mu\nu\rho} \big(C_2 \hat{R}_{\mu\nu}{}^{\sigma}{}_\sigma A_\rho
%%     + C_3 A_\mu\partial_\nu A_\rho\big)\\%\nonumber
%%     & + C_4 T^{\mu\nu}\hat{\nabla}_\mu A_\nu
%%     + C_5 T^{\mu\nu} A_\mu A_\nu
%%     + C_6 \det(T)\\%\nonumber
%%     & + C_7 \epsilon^{\mu\nu\lambda}\big( \hat{\Gamma}^{\sigma}{}_{\mu\rho}\partial_\nu\hat{\Gamma}^\rho{}_{\lambda\sigma}+\frac{2}{3}\hat{\Gamma}^\tau{}_{\mu\rho}\hat{\Gamma}^\rho{}_{\nu\sigma}\hat{\Gamma}^\sigma{}_{\lambda\tau}\big)
%%     \bigg)
%%   \end{split}
%% \end{equation}
%% where  the Chern-Simons term has been included, as it is also invariant.

\section{4D metricless (and torsionful) action}

Following the precepts  already learnt, one starts  by defining a decomposition for the full connection field 
\begin{equation}
  \Gamma^\mu{}_{\rho\sigma} = \hat{\Gamma}^\mu{}_{\rho\sigma} + T^\mu{}_{\rho\sigma} = \hat{\Gamma}^\mu{}_{\rho\sigma} + \epsilon_{\rho\sigma\lambda\kappa}T^{\mu,\lambda\kappa}+A_{[\rho}\delta^\mu_{\nu]},
\end{equation}
where $\hat{\Gamma}^\mu{}_{\rho\sigma}$ denotes a symmetric connection, $A_\mu$ is a vector field  that gives trace to the antisymmetric part of the full connection, and  $T^{\mu,\lambda\kappa}$ is a (Curtright) field that  defined through the symmetry of its indices: antisymmetric in the last two indices, and it has a cyclic property $T^{\mu,\lambda\kappa}+T^{\lambda,\kappa\mu}+T^{\kappa,\mu\lambda}=0$. In other words that $T^{[\mu,\lambda]\kappa}=\frac{1}{2}T^{\kappa,\lambda\mu}$, just as for  the Riemmann tensor $\hat{R}_{\mu[\nu}{}^\alpha{}_{\lambda]}=\frac{1}{2}\hat{R}_{\lambda\nu}{}^\alpha{}_{\mu}$. Notice that due to its symmetries, the contraction $\epsilon_{\rho\sigma\lambda\kappa}T^{\mu,\lambda\kappa}$ is traceless. Additionally, since no metric is present  the epsilon symbols are not related by lowering or raising their indices, but instead one demands that $\epsilon^{\delta\eta\lambda\kappa}\epsilon_{\mu\nu\rho\sigma}=4!\delta^{\delta}{}_{[\mu}\delta^\eta{}_{\nu}\delta^{\lambda}{}_{\rho} \delta^\kappa{}_{\sigma]}$.

One can  write all the combinations of fields that would presumably be renormalizable with these three independent fields
\begin{equation}
  \begin{split}
    \label{3+1full}
     S[\hat{\Gamma}^{\nu}{}_{\mu\rho},T^{\lambda,\mu\nu},A_{\mu}] =\int \dn{4}{x} & \Bigg( A_1 \hat{ R}_{\mu\nu}{}^{\alpha}{}_\beta \hat{ R}_{\rho\sigma}{}^{\beta}{}_\alpha\epsilon^{\mu\nu\rho\sigma}+A_2 \hat{ R}_{\mu\nu}{}^{\alpha}{}_\alpha \hat{ R}_{\rho\sigma}{}^{\beta}{}_\beta\epsilon^{\mu\nu\rho\sigma}
    \\
    & + \Big(B_1 \hat{ R}_{\mu\nu}{}^{\mu}{}_\rho T^{\nu,\alpha\beta}T^{\rho,\gamma\delta}+B_2 \hat{ R}_{\mu\nu}{}^{\alpha}{}_\rho T^{\beta,\mu\nu}T^{\rho,\gamma\delta}+B_3\hat{ R}_{\mu\nu}{}^{\alpha}{}_\rho T^{\mu,\nu\beta}T^{\rho,\gamma\delta}\Big)\epsilon_{\alpha\beta\gamma\delta}
    \\
    & + B_4\hat{ R}_{\mu\nu}{}^{\alpha}{}_\rho T^{\rho,\mu\nu}A_\alpha+B_5\hat{ R}_{\mu\nu}{}^{\rho}{}_\rho T^{\alpha,\mu\nu}A_\alpha+B_6 \hat{ R}_{\mu\nu}{}^{\mu}{}_\rho T^{\rho,\nu\alpha}A_\alpha 
    \\
    & + C_1\hat{ R}_{\mu\nu}{}^{\alpha}{}_\rho \hat{\nabla}_\alpha T^{\rho,\mu\nu}+C_2\hat{ R}_{\mu\nu}{}^{\rho}{}_\rho \hat{\nabla}_\alpha T^{\alpha,\mu\nu} +C_3 \hat{ R}_{\mu\nu}{}^{\mu}{}_\rho \hat{\nabla}_\alpha T^{\rho,\nu\alpha}
    \\
    & +C_4\hat{ R}_{\mu\nu}{}^{\tau}{}_\tau \partial_\rho A_\sigma \epsilon^{\mu\nu\rho\sigma}+D_1T^{\alpha,\mu\nu}T^{\beta,\rho\sigma}\hat{\nabla}_\gamma T^{\lambda,\gamma\kappa}\epsilon_{\mu\nu\beta\lambda}\epsilon_{\alpha\rho\sigma\kappa}
    \\  
    & + D_2T^{\alpha,\mu\nu}T^{\lambda,\beta\gamma}\hat{\nabla}_\lambda T^{\delta,\rho\sigma}\epsilon_{\alpha\beta\gamma\delta}\epsilon_{\mu\nu\rho\sigma}+D_3T^{\mu,\alpha\beta}T^{\lambda,\nu\gamma}\hat{\nabla}_\lambda T^{\delta,\rho\sigma}\epsilon_{\alpha\beta\gamma\delta}\epsilon_{\mu\nu\rho\sigma}
    \\
    & + D_4T^{\lambda,\mu\nu}T^{\kappa,\rho\sigma}\hat{\nabla}_\lambda A_\kappa \epsilon_{\mu\nu\rho\sigma}+D_5T^{\lambda,\mu\nu}A_\lambda\hat{\nabla}_\mu A_\nu +D_6T^{\lambda,\mu\nu}A_\mu\hat{\nabla}_\lambda A_\nu
    \\
    & + D_7A_\lambda  A_\mu\hat{\nabla}_\nu T^{\lambda,\mu\nu} +E_1\hat{\nabla}_\alpha T^{\alpha,\mu\nu}\hat{\nabla}_\beta T^{\beta,\lambda\rho}\epsilon_{\mu\nu\lambda\rho}+E_2\hat{\nabla}_\beta T^{\alpha,\mu\nu}\hat{\nabla}_\alpha T^{\beta,\lambda\rho}\epsilon_{\mu\nu\lambda\rho}
    \\
    & + E_3\hat{\nabla}_\lambda T^{\lambda,\mu\nu}\hat{\nabla}_\mu A_\nu+E_4\hat{\nabla}_\mu T^{\lambda,\mu\nu}\hat{\nabla}_\nu A_\lambda + E_5\partial_\alpha A_\beta\partial_\gamma A_\delta \epsilon^{\alpha\beta\gamma\delta}
    \\
    & + T^{\alpha,\beta\gamma}T^{\delta,\eta\kappa}T^{\lambda,\mu\nu}T^{\rho,\sigma\tau}\Big(\Lambda_1\epsilon_{\beta\gamma\eta\kappa}\epsilon_{\alpha\rho\mu\nu}\epsilon_{\delta\lambda\sigma\tau}+\Lambda_2\epsilon_{\beta\lambda\eta\kappa}\epsilon_{\gamma\rho\mu\nu}\epsilon_{\alpha\delta\sigma\tau}\Big)
    \\
    & + \Lambda_3 T^{\rho,\alpha\beta}T^{\gamma,\mu\nu}T^{\lambda,\sigma\tau}A_\tau \epsilon_{\alpha\beta\gamma\lambda}\epsilon_{\mu\nu\rho\sigma}+\Lambda_4T^{\eta,\alpha\beta}T^{\kappa,\gamma\delta}A_\eta A_\kappa\epsilon_{\alpha\beta\gamma\delta}\Bigg)
  \end{split}
\end{equation}


\subsection*{3+1 Decomposition  and Degrees of Freedom}

The  analysis of the degrees of freedom of the 2+1 action is pretty straightforward because it clearly corresponds to a gravitational action plus some extra stuff. In 3+1 dimensions there is not a clear  equivalence between the model described by Eq.~\eqref{3+1full} to anything known, an deeper  analysis should be done. Hence, one proposes a decomposition of the fields
\begin{align}
  A_\mu &= \delta_\mu^0 A_0+\delta_\mu^iA_i \\
  \intertext{and }
  T^{\mu,\nu\rho} &= \delta^{\mu}_i\delta^{\nu\rho}_{jk}T^{i,jk}+(\delta^{\mu}_0\delta^{\nu\rho}_{ij}-\delta^{\mu}_i\delta^{\nu\rho}_{j0})a^{[ij]}+\delta^{\mu}_i\delta^{\nu\rho}_{j0}T^{(ij)}+\delta^{\mu}_0\delta^{\nu\rho}_{i0}c^i,
\end{align}
where $\delta^{\mu\nu}_{\lambda\kappa}=\delta^{\mu}_{\lambda}\delta^{\nu}_{\kappa}-\delta^{\mu}_{\kappa}\delta^{\nu}_{\lambda}$, $T^{i,jk}\epsilon_{ijk}=0$, $a^{ij}$ is antisymmetric, $T^{ij}$ is a symmetric tensor and $c^i$ an arbitrary vector. In order to make perturbation theory we will expand around an isotropic and homogeneous solution, as these are characteristic of the observable universe. Thus, in order to find such solution we can substitute $A_\mu=\delta_\mu^0 A_0$ and $T^{\mu,\nu\rho}=\delta^{\mu}_i\delta^{\nu\rho}_{j0}T^{ij}$; for a spacially flat metric type, with $k=0$, we would choose $T^{ij}=\delta^{ij}T(t)$. For this particular choice we wish to remark that the contraction of torsion terms
\begin{align}
  T_{\mu}{}^{\sigma}{}_\rho T_{\nu}{}^{\rho}{}_\sigma &= (\epsilon_{\mu\rho\lambda\kappa}T^{\sigma,\lambda\kappa}+A_{[\mu}\delta^\sigma_{\rho]})(\epsilon_{\nu\sigma\delta\eta}T^{\rho,\delta\eta}+A_{[\nu}\delta^\rho_{\sigma]}) \notag \\
  &= \delta^0_\mu\delta^0_\nu\frac{3}{4} A_0^2-24\det(T^{ij})\delta^i_\mu\delta^j_\nu T_{ij},
\end{align}
where $T_{ij}$ is defined as the inverse of $T^{ij}$. Being a covariant tensor, $T_{\mu}{}^{\sigma}{}_\rho T_{\nu}{}^{\rho}{}_\sigma$ can play the role of the metric of the {\it spacetime}, defined as a constructed field $g_{\mu\nu}:= T_{\mu}{}^{\sigma}{}_\rho T_{\nu}{}^{\rho}{}_\sigma $. Similarly, we can define another metric-like tensor, 
\begin{equation}
  \tilde g^{\lambda\kappa}:= T_{\mu}{}^{\lambda}{}_\nu T_{\rho}{}^{\kappa}{}_\sigma\epsilon^{\mu\nu\rho\sigma}=-8\delta^\lambda_i\delta^\kappa_jA_0T^{ij},
\end{equation}
but we can see that this would not be a four dimensional metric but only the space part of it.

Anyhow, we can now do perturbation theory expanding up to second order the fields around the solution $T^{\mu,\nu\lambda} = \delta^{\mu}_i \delta_{i0}^{\nu\lambda}T + t^{\mu,\nu\lambda}$ and $A_\mu=\delta_\mu^0A_0+a_\mu$ and in the simplest case, where $A_0$ and $T$ are constants in time we can also set $\hat{\Gamma}_\mu{}^\lambda{}_\nu=\gamma_\mu{}^\lambda{}_\nu$, where the lowcase fields are small perturbations. The expansion is required to be performed around a minimum of the action, therefore we will set linear terms to zero. 
\begin{equation}
  \label{3+1fullPerturbed}
   S_0[\hat{\Gamma}_{\mu}{}^{\nu}{}_\rho,T^{\lambda,\mu\nu},A_\mu]=\int\dn{4}{x}\Bigg(-3!A_0T^3\Lambda_3\Bigg),
\end{equation}

\begin{equation}
  \begin{split}
    \label{3+1full1stOrder}
    S_1[\hat{\Gamma}_{\mu}{}^{\nu}{}_\rho,T^{\lambda,\mu\nu},A_\mu] = \int\dn{4}{x} &\Bigg(2(-4B_2+B_3)T^2\partial_{[i}\gamma_{0]}{}^k{}_j\epsilon_{0ijk} \\
    & + 2 A_0 T \big((B_4-B_6)\partial_{[i}\gamma_{0]}{}^0{}_i+B_6\partial_{[i}\gamma_{j]}{}^j{}_i\big)\\
    & + 8 T^3 \big(-3D_1\gamma_0{}^0{}_0+(-2D_1+D_2+D_3)\gamma_i{}^i{}_0\big)\\
    & - T A_0^2(D_6+D_7)\gamma_i{}^0{}_i -12\Lambda_3( 2t^{k,k0}+a_0)\Bigg).
  \end{split}
\end{equation}
Therefore, we must set $B_3=4B_2$, $B_4=0=B_6$, $D_1=0$, $D_3=-D_2$,  $D_7=-D_6$ and $\Lambda_3=0$ in order to be sure that the homogeneous and isotropic field configuration minimizes the action. 

The second order perturbative expansion  is

\section*{FALTA}

\begin{itemize}
\item tal vez $a^{ij}$ y $c^i$ sean campo el\'ectrico y magn\'etico

\item study  whether the spacetime solution can arise as a symmetry breaking from thermal fluctuations around $T=0$

\item estudiar la posibilidad de obtener Ho\v{r}ava-Lifshitz gravity en algun limite

\item hacer teoria de perturbaciones alrededor de la solucion constante

\item calcular el limite Newtoniano

\item I am studying how to do the d.o.f. analysis.
\item check the most general action in 4d.
\end{itemize}

\nocite{Tucker:1996sx,Horava:2009uw,Lu:2009em,Gibbs:1995gj,WheelerPre,Peeters:2007wn,peeters2007symbolic,Peeters2007550,sage}

\bibliographystyle{aipauth4-1}
\bibliography{References.bib}

\end{document}

%% In the case that all $D_\pi=0$ we are left with an action that depends 
%% directly on a combination of the Riemann and Torsion tensor. 
%% Determination of the equations of motion is straightforward, and we obtain 
%% a combination of $\nabla_\mu T_{\nu}{}^\sigma{}_{\lambda}$ and 
%% $R_{\mu\nu}{}^\sigma{}_{\lambda}$. However, due to the complicated 
%% combination of invariant tensors, we can solve these equations by setting
%%  $\nabla_\mu T_{\nu}{}^\sigma{}_{\lambda}=0$ and 
%% $R_{\mu\nu}{}^\sigma{}_{\lambda}=0$. 

%% Thus, it seems appropriate to decompose it in terms of $
%% \Gamma_{\nu}{}^\lambda{}_{\rho}=\hat{\Gamma}_{\nu}{}^\lambda{}_{\rho}+\hat{ T}_{\nu}{}^\lambda{}_{\rho}+A_{[\nu}\delta^\lambda_{\rho]}$. Then,
%% \ba
%% R_{\mu\nu}{}^\lambda{}_\rho&=&2\hat{\nabla}_{[\mu}\hat{\Gamma}_{\nu]}{}^\lambda{}_{\rho}+2\hat{\nabla}_{[\mu}\hat{ T}_{\nu]}{}^\lambda{}_{\rho}-2\hat{ T}_{\sigma}{}^\lambda{}_{[\mu}\hat{ T}_{\nu]}{}^\sigma{}_{\rho}+\\ \nonumber&&+\hat{\nabla}_{[\mu}A_{\nu]}\delta^\lambda_\rho+\delta^\lambda_{[\mu}\hat{\nabla}_{\nu]}A_\rho-\hat{ T}_\mu{}^\lambda{}_{\nu}A_\rho-\delta^\lambda_{[\mu}\hat{ T}_{\nu]}{}^\sigma{}_{\rho}A_\sigma +\frac{1}{2}\delta^\lambda_{[\mu}A_{\nu]}A_\rho
%% \ea

%% {\bf esto se ve tan feo que creo que es preferible trabajar con la torsion sin hacer la descomposicion de la torsion en $\hat{ T} +A$}

%% Thus, we decompose the connection in terms of its symmetric and antisymmetric parts $
%% \Gamma_{\nu}{}^\lambda{}_{\rho}=\hat{\Gamma}_{\nu}{}^\lambda{}_{\rho}+\frac{1}{2} T_{\nu}{}^\lambda{}_{\rho}$
%% \be
%% R_{\mu\nu}{}^\lambda{}_\
